---
title: "R Notebook"
output: html_notebook
---

```{r}
# Import needed libraries
library(readr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(RCurl)
library(forcats)
library(broom)
library(rvest)
library(Metrics)

# Import turnover data
turnover_data <- read_csv("https://raw.githubusercontent.com/opencampus-sh/einfuehrung-in-data-science-und-ml/main/umsatzdaten_gekuerzt.csv", show_col_types = FALSE)

#import Kieler Woche data
kiwo <- read_csv("https://raw.githubusercontent.com/opencampus-sh/einfuehrung-in-data-science-und-ml/main/kiwo.csv", show_col_types = FALSE)

#import weather data
weather <- read_csv("https://raw.githubusercontent.com/opencampus-sh/einfuehrung-in-data-science-und-ml/main/wetter.csv", show_col_types = FALSE)

#import extended weather Data from "Deutscher Wetterdienst"
weatherDWD <- read_delim("https://raw.githubusercontent.com/g3r1t/Umsatzdaten_Dump/main/Klimadaten-DWD-20130701-20191231.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, show_col_types = FALSE)

#import monthly overnight stay and arrival data for Kiel
tourism <- read_csv("https://raw.githubusercontent.com/g3r1t/Umsatzdaten_Dump/main/Ankuenfte%20und%20Uebernachtungen.csv", show_col_types = FALSE)

#import holidays
Feiertage <- read_csv("https://raw.githubusercontent.com/g3r1t/Umsatzdaten_Dump/main/Feiertage.csv", 
                      show_col_types = FALSE)

#import school holidays
Ferien <- read_csv("https://raw.githubusercontent.com/g3r1t/Umsatzdaten_Dump/main/Ferien.csv", 
                    show_col_types = FALSE)

#left join to keep all data points from turnoverdata and drop the redundant ones from kiwo
full_Dataset <- left_join(turnover_data, kiwo, by = "Datum")

#replaces all NAs of days where there is no KiWo with 0 and makes it a categorical boolean variable
full_Dataset$KielerWoche[is.na(full_Dataset$KielerWoche)] <- 0

#again left join to only keep data points with turnover
full_Dataset <- left_join(full_Dataset, weather, by = "Datum")

#change "Datum"-column of weatherDWD to date-format
weatherDWD$Datum <- as.Date(weatherDWD$Datum, "%d.%m.%Y")

#leftjoin weather DWD and full_Dataset
full_Dataset <- left_join(full_Dataset, weatherDWD, by="Datum")

#replace NAs from weather with data from DWD
full_Dataset$Temperatur.x[is.na(full_Dataset$Temperatur.x)] <- full_Dataset$Temperatur.y[is.na(full_Dataset$Temperatur.x)]

full_Dataset$Windgeschwindigkeit.x[is.na(full_Dataset$Windgeschwindigkeit.x)] <- full_Dataset$Windgeschwindigkeit.y[is.na(full_Dataset$Windgeschwindigkeit.x)]

full_Dataset$Bewoelkung[is.na(full_Dataset$Bewoelkung)] <- full_Dataset$Bedeckungsgrad[is.na(full_Dataset$Bewoelkung)]

#Create Column Monatscode to left_join tourism data with full_Dataset
for (e in as.numeric(row.names(full_Dataset))) {
  full_Dataset$Monatscode[e] <- paste(
  (year(full_Dataset$Datum[e])-2000),
  formatC(month(full_Dataset$Datum[e]), width = 2, format = "d", flag = "0"), sep = "")
  }
full_Dataset$Monatscode <- as.numeric(full_Dataset$Monatscode)
full_Dataset <- left_join(full_Dataset, tourism, by="Monatscode")

#join holidays/Feiertage with dataset
full_Dataset <- left_join(full_Dataset, Feiertage, by = "Datum")
full_Dataset[c(26:41)][is.na(full_Dataset[c(26:41)])] <- 0 #set all NAs to zero for boolean variable

#join school holidays/Ferien with dataset
full_Dataset <- left_join(full_Dataset, Ferien, by = "Datum", suffix = c(".Feiertage", ".Ferien"))
full_Dataset[c(42:57)][is.na(full_Dataset[c(42:57)])] <- 0 #set all NAs to zero for boolean variable
```


```{r}
#import and join weather codes
webpage <- read_html(paste("http://www.seewetter-kiel.de/seewetter/daten_symbole.htm")) #read html page
weather_codes <- html_table(webpage, trim = TRUE)#turn all html-tables into a list of table-DFs
weather_codes <- weather_codes[[1]]# select first element of list of DFs

weather_codes <- weather_codes[-1]#drop first column

colnames(weather_codes)[1] <- "Wettercode"#rename first column to for left_join with full_Dataset

full_Dataset <- left_join(full_Dataset, weather_codes, by = "Wettercode")

#Group Weathercodes
full_Dataset$Wettergruppe[full_Dataset$Wettercode <= 3] <- "?Bewölkung"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 4 & full_Dataset$Wettercode <= 9] <- "Dunst, Rauch, Staub oder Sand"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 10 & full_Dataset$Wettercode <= 19] <- "Trockenereignisse"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 20 & full_Dataset$Wettercode <= 29] <- "Ereignisse der letzten Stunde, aber nicht zur Beobachtungszeit"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 30 & full_Dataset$Wettercode <= 39] <- "Staubsturm, Sandsturm, Schneefegen oder -treiben"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 40 & full_Dataset$Wettercode <= 49] <- "Nebel oder Eisnebel"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 50 & full_Dataset$Wettercode <= 59] <- "Sprühregen"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 60 & full_Dataset$Wettercode <= 69] <- "Regen"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 70 & full_Dataset$Wettercode <= 79] <- "Schnee"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 80 & full_Dataset$Wettercode <= 89] <- "Schauer"
full_Dataset$Wettergruppe[full_Dataset$Wettercode >= 90 & full_Dataset$Wettercode <= 99] <- "Gewitter"
full_Dataset$Wettergruppe[is.na(full_Dataset$Wettercode)] <- "NA"

#Convert Dates to weekdays
full_Dataset$Wochentag <- weekdays(full_Dataset$Datum)

# Create vector of product names
Warenname <- c("Brot", "Brötchen", "Crossaint", "Konditorei", "Kuchen", "Saisonbrot")

#Convert Warengruppe to Warenname
#for each "Warengruppe"-number at position "[e]" a new row entry in column "Warenname" is created. This works bcs the index of the "Warengruppenname" in the vector corresponds with the number of "Warengruppe".
for (e in as.numeric(row.names(full_Dataset))) #enumerate Dataframe, since row.names returns a string
  full_Dataset$Warenname[e] <- Warenname[as.numeric(full_Dataset$Warengruppe[e])]

#Add month variable
full_Dataset$Monat <- month(full_Dataset$Datum, label = TRUE)

#Add calendar week variable
full_Dataset$KalenderWoche <- strftime(full_Dataset$Datum, format = "%V")

#Add Wochenende create boolean vectors for all days == Samstag | Sonntag and for all day != Samstag & Sonntag
full_Dataset$Wochenende[full_Dataset$Wochentag == "Samstag" | full_Dataset$Wochentag == "Sonntag"] <- 1
full_Dataset$Wochenende[full_Dataset$Wochentag != "Samstag" & full_Dataset$Wochentag != "Sonntag"] <- 0

#Add windchill Temp(https://de.wikipedia.org/wiki/Windchill)
full_Dataset$Windchill <- 13.12 + 0.6215 * full_Dataset$Temperatur.x + (0.3965 * full_Dataset$Temperatur.x - 11.37) * ((full_Dataset$Windgeschwindigkeit.x*3600)/1000)**0.16

write.csv(full_Dataset, "full_Dataset.csv", row.names = FALSE)
```


ToDo:
-VariablenTyp checken -> z.B. KW=factor Variable
-Temperatur Chaos lösen, scheinbar ist Temperatur und Windgeschwindigkeit aus weather=max. Temperatur und Windgeschwindigkeit vom DWD aber Durschnitt
-Saisonbrot gibt es nur von Oktober bis Januar
-Feiertage anders einbinden, jede Bundesland in einer Spalte ist..... nervig und schadet den Kosten
-fix singularities